Moral Foundations of U.S. Political News Organizations

### 2019-03-17 pilot_manuscript.Rmd
wrote about MFT criticisms and validation of MFQ

@@

>>>>

Moral Foundations Theory has received criticism on the grounds that its assumptions regarding moral intuitions have little empirical basis. @Suhler2011 list several potential weaknesses of MFT that they argue might threaten the theory's validity. First, the authors challenge @Haidt2007's claims the moral intuitions represented by MFT are innate and modular. @Suhler2011 claim that advances in biological sciences (embryology and microbiology, specifically) make it more difficult for researchers to claim any one trait is either innate or learned through experience. Rather, behaviors likely result from interactions between genetics and experience [@Suhler2011]. According to the authors, without solid data supporting the innateness of moral foundations, Haidt and Graham have little from which to make such a claim. Similarly, @Haidt2007 rely on evidence authored by evolutionary psychologists to make a "strong modularity claim" [@Suhler2011]. However, as with innateness, there is little neurobiological evidence to support modularity.

@Suhler2011 also criticized the content and taxonomy of the five foundations. The authors criticize @Haidt2007's omissions of additional foundations, including *industry* and *modesty*, claiming these concepts are moralized in many societies worldwide. Likewise, the authors question whether or not the foundations are sufficiently distinct as to stand as their own foundation. For example, @Suhler2011 posit that *ingroup/loyalty* is merely a group-focused version of *harm/care*. Work by @Graham2011 might serve to rebut this criticism, as the researchers found their original five-factor structure seemed to best fit the data when validating the Moral Foundations Questionnaire. Finally, @Suhler2011 point out that particular concepts related to a foundation, including "anger" in *fairness/reciprocity* and "deception" in *ingroup/loyalty*, could be ascribed to any of the other foundations as well. In other words, it becomes difficult to recognize a particular concept as indidcative of any one foundation when, in theory, they could be applied to all five [@Haidt2007; @Suhler2011].

These criticisms of Moral Foundations Theory are valid and should be taken into consideration when conducting research with instruments derived from MFT. The current authors argue these criticisms are especially valid when considered alongside questions regarding the Moral Foundations Dictionary they state herein. However, there exists compelling evidence regarding the validity of Moral Foundations Theory, albeit regarding its application solely through the Moral Foundations Questionnaire. This evidence is discussed within a brief explanation of the Moral Foundations Questionnaire and its relationship to the Moral Fondations Dictionary.

@@

The Moral Foundations Questionnaire has been validated by multiple researchers. Likewise, its five-factor structure has been demonstrated to fit data in multiple countries [@Graham2011; @Davies2014]. In their article introducing the 30-item MFQ, @Graham2011 conducted an exploratory factor analysis (EFA) and found that a five-factor model fit better than a one, two, or three-factor model. @Davies2014 conducted a confirmatory factor analysis (CFA) of the MFQ with a sample from New Zealand and likewise found the five-factor model provided the best fit. While @Davies2014 concede the US and New Zealand share many similarities as Western nations, which could raise questions regarding the validity of the MFQ in non-Western nations. However, there are striking differences between the two countries, including the lack of a two-party political system in New Zealand, that provide grounds for claiming the MFQ generalizes beyond the United States. Furthermore, @Graham2011 claim to find a "reasonable" degree of generalizability for the MFQ across participants from many different regions in the world. These two bodies of work also provide the best available evidence that the five moral foundations are sufficiently distinct from one another, though broader criticisms of MFT raised by @Shuler2011 should still be taken into account in studies involving the theory.
